---
title: "Bellabeat Capstone Report"
author: "Paola Velleda Betemps"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, fig.height = 5, dpi = 150)
library(tidyverse); library(lubridate); library(janitor); library(here); library(skimr); library(scales)
Sys.setlocale("LC_TIME","C")
# Build plots p1..p6 used below
source(here::here("03_analysis","01_explore.R"))
```

## Introduction
This report presents an analysis of Fitbit daily activity and sleep data to provide insights for Bellabeat’s smart wellness products.

## Data & KPIs
```{r}
clean_path <- file.path(here::here("02_clean"), "bellabeat_daily_clean.csv")
daily <- readr::read_csv(clean_path) |> janitor::clean_names()
daily <- daily |> mutate(dow = lubridate::wday(date, label=TRUE, abbr=TRUE, week_start=1),
                         sleep_hours = total_minutes_asleep/60)

kpi <- daily |>
  summarise(users = dplyr::n_distinct(id),
            rows  = dplyr::n(),
            steps_median = median(total_steps, na.rm=TRUE),
            steps_mean   = mean(total_steps, na.rm=TRUE),
            sleep_median_h = median(sleep_hours, na.rm=TRUE),
            sleep_mean_h   = mean(sleep_hours, na.rm=TRUE),
            corr_steps_sleep = cor(total_steps, total_minutes_asleep, use="complete.obs"))
kpi
```

## Visual Insights

### 1) Distribution of Daily Steps
```{r}
p1
```

### 2) Sleep (hours) vs. Steps
```{r}
p2
```

### 3) Average Steps by Weekday
```{r}
p3
```

### 4) Average Sleep by Weekday
```{r}
p4
```

### 5) Median Steps by ISO Week
```{r}
p5
```

### 6) Top 10 Users by Average Steps
```{r}
p6
```

## Recommendations
- Set weekday-specific goals to address variability.
- Provide sleep coaching nudges after unusually high/low activity days.
- Send weekly trend summaries to sustain engagement.
